---
title: "COVID-19 Project: LOESS Curves for 29 Genes and 16 Brain Regions"
author: "Rachel Kwan and Jonathan Zaslavsky"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include = FALSE}
library(knitr) # For formatting the document
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
```

## Relevant Packages
```{r}
library(tidyverse) # For ggplot2, dplyr
library(here) # To read in data from directory
library(reshape2) # For melt()
library(R.devices) # Ease of use for the pdf function
library(ggforce) # For the facet_wrap_paginate()
library(pdftools) # For pdf_combine()
```

## Import and Prepare Dataset
```{r}
# Load full dataset
all_expression <- read.csv(here("Data", "Kang Data 17528 genes 1340 cases.csv")) 


###===== Subset of 29 Genes =====###

# List of 12 SARS-CoV-2 receptors:
# ANPEP, ENPEP, ACE2, DPP4, SCARB1, AXL, NRP1, NRP2, CLEC4M, CLEC4G, CD209, CEACAM1
# List of 4 SARS-CoV-2 proteases:
# TMPRSS2, FURIN, ADAM17, CTSL
# List of 13 SARS-CoV-2 integrins:
# ITGB3, ITGA5, ITGB1, ITGB6, ITGAV, ITGB8, ITGB2, ITGA2B, ITGAM, ITGAL, ITGA3, ITGA8, ITGB5

# Create a vector containing all of the names of the 29 genes of interest,
# and sort into alphabetical order.
genes <- c('ANPEP', 'ENPEP', 'ACE2', 'DPP4', 'SCARB1', 'AXL', 'NRP1', 'NRP2', 'CLEC4M', 
           'CLEC4G', 'CD209', 'CEACAM1', 'TMPRSS2', 'FURIN', 'ADAM17', 'CTSL1', 'ITGB3', 
           'ITGA5', 'ITGB1', 'ITGB6', 'ITGAV', 'ITGB8', 'ITGB2', 'ITGA2B', 'ITGAM', 
           'ITGAL', 'ITGA3', 'ITGA8', 'ITGB5') %>%
  sort()

# Define a vector containing the 16 brain regions of interest (ordered alphabetically).
regions_16 <- c('A1C', 'AMY', 'CBC', 'DFC', 'HIP', 'IPC', 'ITC', 'M1C', 'MD', 'MFC', 'OFC',
                'S1C', 'STC', 'STR', 'V1C', 'VFC')

# Select desired columns from original full data frame, then filter to keep only 
# desired regions.
all_expression <- all_expression %>%
  select(Sample, Region, Years, genes) %>%
  filter(Region %in% regions_16)

# Add 40/52 to the 'Years' column to get rid of negative values.
all_expression$Years <- all_expression[, 3] + 40/52
```

# 1. LOESS Stamp Collections

## 1.1. Z-Scored Data

### Prepare data frame for LOESS Curves (Z-scored, Non-Uniform Age Values)
```{r z-score the data}
# Specify order of regions to display in LOESS curves pdf
new_regions_order <- c('VFC', 'DFC', 'MFC', 'OFC', 'M1C', 'S1C', 'IPC', 'STC',
                       'A1C', 'ITC', 'V1C', 'MD', 'AMY', 'HIP', 'STR', 'CBC')

covid.gene.exp.numeric <- all_expression %>%
  #Make sure all expression values are numeric in R
  mutate_at(genes, as.numeric)

covid.gene.exp.zscore <- covid.gene.exp.numeric %>%
  # Group the data by region
  group_by(Region) %>% 
  # 'Nest' the data (i.e. build data frames for each region inside of the data frame)
  nest() %>%
  # z-score the data separately for each gene and region
  mutate(data = map(.x = data, .f = ~.x %>% mutate_at(genes, scale))) %>%
  # Unnest the data
  unnest(cols = c(data))

# Reorder data frame to match desired order
covid.gene.exp.zscore.ordered <- covid.gene.exp.zscore[order(factor(covid.gene.exp.zscore$Region, 
                                                                    levels = new_regions_order)),]

# Get long version of z-scored data
covid.gene.exp.zscore.long <- covid.gene.exp.zscore.ordered %>%
  pivot_longer(cols = all_of(genes),
               names_to = "Gene",
               values_to = "Expression")
```

### Fit LOESS Curves for Z-scored Data
```{r fit loess curves for z-scored data}
covid.gene.exp.zscore.loess <- covid.gene.exp.zscore %>%
  # Use the melt function to put the data in long format
  melt(id.vars = c("Sample", "Region", "Years"),
       variable.name = "Gene",
       value.name = "Expression") %>%
  # Group by region and Gene, then nest the data
  group_by(Region, Gene) %>%
  nest() %>%
  # Fit a LOESS model for each gene and region
  mutate(loess.model = map(.x = data, ~loess(data = .x, formula = Expression ~ log2(Years)))) %>%
  # Find the min and max age value for each curve
  mutate(min = map(.x = data, .f = ~min(log2(.x$Years))),
         max = map(.x = data, .f = ~max(log2(.x$Years)))) %>%
  unnest(min) %>% unnest(max) %>%
  # Create a vector of 80 log2 age values for each curve
  mutate(age.values = map2(.x = min, .y = max, ~seq(.x, .y, (.y - .x) / 79))) %>%
  # Extract the fitted LOESS expression values along this curve
  mutate(fitted.values = map2(.x = loess.model, .y = age.values,
                              ~data.frame("Years" = 2^(.y), 
                                          "Expression" = predict(.x, newdata = .y))))

# Create a data frame including only the fitted values of the LOESS curves of
# the z-scored data
covid.gene.exp.zscore.loess.fitted = covid.gene.exp.zscore.loess %>%
  # Select necessary columns
  select(Region, Gene, fitted.values) %>%
  # Unnest the fitted values
  unnest(cols = c(fitted.values))
```

```{r save csv}
# Save the z-scored LOESS fitted value data as a csv
write.csv(covid.gene.exp.zscore.loess.fitted, 
          here::here("Data", 
                     "COVID Data - Zscored - Fitted LOESS Values for 29 Genes and 16 Regions.csv"),
          row.names = FALSE)
```

### Plot a Test LOESS Curve

Test LOESS Curve for Z-scored Data
```{r}
test.raw = covid.gene.exp.zscore.long %>% 
  # Filter for only DPP4 expression values from the V1C
  filter(Gene == "DPP4", Region == "A1C")

ggplot(test.raw, aes(x = Years, y = Expression)) +
  # Add in shading in the background to separate pre-natal, infancy, childhood, and adulthood
  annotate("rect", xmin = 0, xmax = 40/52, ymin = -Inf, ymax = Inf,
           alpha = 0.2, fill = "grey40") +
  annotate("rect", xmin = 40/52, xmax = 92/52, ymin = -Inf, ymax = Inf,
           alpha = 0.2, fill = "grey60") +
  annotate("rect", xmin = 92/52, xmax = 40/52 + 18, ymin = -Inf, ymax = Inf,
           alpha = 0.2, fill = "grey80") +
  # Add in a dashed line to separate pre and post-natal
  geom_vline(aes(xintercept = 40/52), linetype = "dashed") +
  # Add in the LOESS curves
  geom_smooth(method = "loess", span = 0.75,
              method.args = list(family = 'gaussian', degree = 2)) +
  # Add points
  geom_point(size = 1) +
  # Scale the x-axis by log2 and overwrite the labels
  scale_x_continuous(trans = "log2", 
                     breaks = c(0, 10/52, 20/52, 40/52, 40/52 + 2/12,
                                40/52 + 6/12, 92/52, 40/52 + 2, 40/52 + 4,
                                40/52 + 8, 40/52 + 16, 40/52 + 32, 
                                40/52 + 64, 40/52 + 90),
                     labels = c("0 PCW", "10 PCW", "20 PCW", "0", "2 M",
                                "6 M", "1 Y", "2 Y", "4 Y", "8 Y", "16 Y",
                                "32 Y", "64 Y", "90 Y")) +
  # Separate the data into 16 panels for each region
  facet_wrap(~Region + Gene) +
  # Label the axes
  labs(x = "Age", y = "Lifespan Trajectory of Gene Expression") +
  # Change up the theme for aesthetics and clarity
  theme_bw() +
  theme(axis.text.x = element_text(angle= 270, size = 7),
        axis.text.y = element_text(size = 7),
        axis.title.y = element_text(size = 9),
        panel.grid = element_blank(),
        strip.text.x = element_text(face = "bold"),
        strip.background = element_rect(fill = "grey75"))
```

### Plot the LOESS Curves and Save As PDFs

#### PDF of LOESS Curves (Z-scored) for 29 Genes by Region 
```{r message=FALSE, warning=FALSE}
# For order of plots in pdf
covid.gene.exp.zscore.long$Region <- factor(covid.gene.exp.zscore.long$Region,
                                            levels = new_regions_order)

# Create the plots and store them as loess.raw.zscore.stamp.collection
loess.raw.zscore.stamp.collection = ggplot(covid.gene.exp.zscore.long, 
                                           aes(x = Years, y = Expression)) +
  # Add in shading in the background to separate pre-natal, infancy, childhood, 
  # and adulthood
  annotate("rect", xmin = 0, xmax = 40/52, ymin = -Inf, ymax = Inf,
           alpha = 0.2, fill = "grey40") +
  annotate("rect", xmin = 40/52, xmax = 92/52, ymin = -Inf, ymax = Inf,
           alpha = 0.2, fill = "grey60") +
  annotate("rect", xmin = 92/52, xmax = 40/52 + 18, ymin = -Inf, ymax = Inf,
           alpha = 0.2, fill = "grey80") +
  # Add in a dashed line to separate pre and post-natal
  geom_vline(aes(xintercept = 40/52), linetype = "dashed") +
  # Add in the LOESS curves
  geom_smooth(method = "loess", span = 0.75,
              method.args = list(family = 'gaussian', degree = 2)) +
  # Add points
  #geom_point(size = 1) +
  # Scale the x-axis by log2 and overwrite the labels
  scale_x_continuous(trans = "log2", 
                     breaks = c(0, 10/52, 20/52, 40/52, 40/52 + 2/12,
                                40/52 + 6/12, 92/52,40/52 + 2, 40/52 + 4,
                                40/52 + 8, 40/52 + 16, 40/52 + 32,
                                40/52 + 64, 40/52 + 90),
                     labels = c("0 PCW", "10 PCW", "20 PCW", "0", "2 M",
                                "6 M", "1 Y", "2 Y", "4 Y", "8 Y", "16 Y",
                                "32 Y", "64 Y", "90 Y")) +
  # Separate the data into 16 panels (regions) for each gene
  facet_wrap(~Gene + Region) +
  # Label the axes
  labs(x = "Age", y = "Lifespan Trajectory of Gene Expression") +
  # Change up the theme for aesthetics and clarity
  theme_bw() +
  theme(axis.text.x = element_text(angle= 270, size = 7),
        axis.text.y = element_text(size = 7),
        axis.title.y = element_text(size = 9),
        panel.grid = element_blank(),
        strip.text.x = element_text(face = "bold"),
        strip.background = element_rect(fill = "grey75"))

# Determine how many pages we will need
num_pages = length(genes)
```

```{r eval=FALSE, include=FALSE}
# Save the plots as a pdf
devEval(type = "pdf",
        path = here("Exploration"),
        name = "LOESS Curves of Z Scored Expression for all 29 Genes By Region with SE no points - wide",
        height = 9, 
        width = 12,
        expr = {
          for (j in 1:num_pages) {
            # Plot LOESS curve for each and facet wrap by gene
              print(loess.raw.zscore.stamp.collection + 
                      facet_wrap_paginate(vars(Gene, Region), 
                                          ncol = 4, nrow = 4, 
                                          page = j, scales = "free_y"))
            }
        })
```

#### PDF of LOESS Curves (Z-scored) for 16 Regions by Gene 
```{r}
# Use for loop to create a separate PDF for each region, each PDF contains
# 29 LOESS plots for each gene. This will result in 16 PDFs which will later
# be merged together into one PDF.
for (i in new_regions_order) {
  
  # Create subset containing loess data for region i only
  subset <- subset(covid.gene.exp.zscore.long, Region == i)
  
  # Create the plots and store them as loess.raw.zscore.stamp.collection
  loess.raw.zscore.stamp.collection <- ggplot(subset, 
                                              aes(x = Years, y = Expression)) +
    # Add in shading in the background to separate pre-natal, infancy, childhood, 
    # and adulthood
    annotate("rect", xmin = 0, xmax = 40/52, ymin = -Inf, ymax = Inf,
             alpha = 0.2, fill = "grey40") +
    annotate("rect", xmin = 40/52, xmax = 92/52, ymin = -Inf, ymax = Inf,
             alpha = 0.2, fill = "grey60") +
    annotate("rect", xmin = 92/52, xmax = 40/52 + 18, ymin = -Inf, ymax = Inf,
             alpha = 0.2, fill = "grey80") +
    # Add in a dashed line to separate pre and post-natal
    geom_vline(aes(xintercept = 40/52), linetype = "dashed") +
    # Add in the LOESS curves
    geom_smooth(method = "loess", span = 0.7091,
                method.args = list(family = 'gaussian', degree = 2)) +
    # Add points
    #geom_point(size = 1) +
    # Scale the x-axis by log2 and overwrite the labels
    scale_x_continuous(trans = "log2", 
                       breaks = c(0, 10/52, 20/52, 40/52, 40/52 + 2/12,
                                  40/52 + 6/12, 92/52, 40/52 + 2, 40/52 + 4,
                                  40/52 + 8, 40/52 + 16, 40/52 + 32,
                                  40/52 + 64, 40/52 + 90),
                       labels = c("0 PCW", "10 PCW", "20 PCW", "0", "2 M",
                                  "6 M", "1 Y", "2 Y", "4 Y", "8 Y", "16 Y",
                                  "32 Y", "64 Y", "90 Y")) +
    # Separate the data into 29 panels (genes) for each region
    facet_wrap(~Region + Gene) +
    # Label the axes
    labs(x = "Age", y = "Lifespan Trajectory of Gene Expression") +
    # Change up the theme for aesthetics and clarity
    theme_bw() +
    theme(axis.text.x = element_text(angle = 270, size = 7),
          axis.text.y = element_text(size = 7),
          axis.title.y = element_text(size = 9),
          panel.grid = element_blank(),
          strip.text.x = element_text(face = "bold"),
          strip.background = element_rect(fill = "grey75")) 
  
  # For each region, we need 2 pages
  # 16 genes on first page, 13 genes on second page
  num_pages2 <- 2
  
  # Save the plots as a pdf
  devEval(type = "pdf",
          path = here("Exploration"),
          name = paste0("LOESS Curves of Z-scored Expression for ", i, " by Gene with SE no points"),
          height = 9, 
          width = 12,
          expr = {
            for (j in 1:num_pages2) {
              # Plot LOESS curve for each region and facet wrap by gene
                print(loess.raw.zscore.stamp.collection + 
                      facet_wrap_paginate(~Region + Gene,
                                          ncol = 4,
                                          nrow = 4,
                                          page = j,
                                          scales = "free_y"))
            }
           
          })

}
```

```{r, merge pdfs into one}
# Merge all 16 PDFs into one — paste path of each pdf and specify output path
pdf_combine(
  c("/Users/rachelkwan/Documents/Lab/summer_2021/Exploration/LOESS Curves of Z-scored Expression for VFC by Gene with SE no points.pdf",
    "/Users/rachelkwan/Documents/Lab/summer_2021/Exploration/LOESS Curves of Z-scored Expression for DFC by Gene with SE no points.pdf",
    "/Users/rachelkwan/Documents/Lab/summer_2021/Exploration/LOESS Curves of Z-scored Expression for MFC by Gene with SE no points.pdf",
   "/Users/rachelkwan/Documents/Lab/summer_2021/Exploration/LOESS Curves of Z-scored Expression for OFC by Gene with SE no points.pdf",
   "/Users/rachelkwan/Documents/Lab/summer_2021/Exploration/LOESS Curves of Z-scored Expression for M1C by Gene with SE no points.pdf",
   "/Users/rachelkwan/Documents/Lab/summer_2021/Exploration/LOESS Curves of Z-scored Expression for S1C by Gene with SE no points.pdf",
   "/Users/rachelkwan/Documents/Lab/summer_2021/Exploration/LOESS Curves of Z-scored Expression for IPC by Gene with SE no points.pdf",
   "/Users/rachelkwan/Documents/Lab/summer_2021/Exploration/LOESS Curves of Z-scored Expression for STC by Gene with SE no points.pdf",
   "/Users/rachelkwan/Documents/Lab/summer_2021/Exploration/LOESS Curves of Z-scored Expression for A1C by Gene with SE no points.pdf",
   "/Users/rachelkwan/Documents/Lab/summer_2021/Exploration/LOESS Curves of Z-scored Expression for ITC by Gene with SE no points.pdf",
   "/Users/rachelkwan/Documents/Lab/summer_2021/Exploration/LOESS Curves of Z-scored Expression for V1C by Gene with SE no points.pdf",
   "/Users/rachelkwan/Documents/Lab/summer_2021/Exploration/LOESS Curves of Z-scored Expression for MD by Gene with SE no points.pdf",
   "/Users/rachelkwan/Documents/Lab/summer_2021/Exploration/LOESS Curves of Z-scored Expression for AMY by Gene with SE no points.pdf",
   "/Users/rachelkwan/Documents/Lab/summer_2021/Exploration/LOESS Curves of Z-scored Expression for HIP by Gene with SE no points.pdf",
   "/Users/rachelkwan/Documents/Lab/summer_2021/Exploration/LOESS Curves of Z-scored Expression for STR by Gene with SE no points.pdf",
   "/Users/rachelkwan/Documents/Lab/summer_2021/Exploration/LOESS Curves of Z-scored Expression for CBC by Gene with SE no points.pdf"),
   output = "/Users/rachelkwan/Documents/Lab/summer_2021/Exploration/LOESS Curves of Z Scored Expression for all 16 Regions by Gene with SE no points - wide.pdf")
```

## 1.2 Non Z-scored Data

### Prepare data frame for LOESS Curves (Raw, Non-Uniform Age Values)
```{r}
covid.gene.exp.numeric <- all_expression %>%
  #Make sure all expression values are numeric in R
  mutate_at(genes, as.numeric)

# Get long version of raw data
covid.gene.exp.long <- covid.gene.exp.numeric %>%
  pivot_longer(cols = all_of(genes),
               names_to = "Gene",
               values_to = "Expression")

covid.gene.exp.loess <- covid.gene.exp.numeric %>%
  # Use the melt function to put the data in long format
  melt(id.vars = c("Sample", "Region", "Years"),
       variable.name = "Gene",
       value.name = "Expression") %>%
  # Group by region and Gene, then nest the data
  group_by(Region, Gene) %>%
  nest() %>%
  # Fit a LOESS model for each gene and region
  mutate(loess.model = map(.x = data, ~loess(data = .x, formula = Expression ~ log2(Years)))) %>%
  # Find the min and max age value for each curve
  mutate(min = map(.x = data, .f = ~min(log2(.x$Years))),
         max = map(.x = data, .f = ~max(log2(.x$Years)))) %>%
  unnest(min) %>% unnest(max) %>%
  # Create a vector of 80 log2 age values for each curve
  mutate(age.values = map2(.x = min, .y = max, ~seq(.x, .y, (.y - .x) / 79))) %>%
  # Extract the fitted LOESS expression values along this curve
  mutate(fitted.values = map2(.x = loess.model, .y = age.values,
                              ~data.frame("Years" = 2^(.y), 
                                          "Expression" = predict(.x, newdata = .y))))

# Create a data frame including only the fitted values of the LOESS curves of
# the z-scored data
covid.gene.exp.loess.fitted = covid.gene.exp.loess %>%
  # Select necessary columns
  select(Region, Gene, fitted.values) %>%
  # Unnest the fitted values
  unnest(cols = c(fitted.values))
```

```{r save csv}
# Save the raw LOESS fitted value data as a csv
write.csv(covid.gene.exp.loess.fitted, 
          here::here("Data", 
                     "COVID Data - Raw - Fitted LOESS Values for 29 Genes and 16 Regions.csv"),
          row.names = FALSE)
```

### Plot the LOESS Curves and Save As PDFs

#### PDF of LOESS Curves (Raw/Non Z-scored) for 29 Genes by Region 
```{r message=FALSE, warning=FALSE}
# Specify order of regions to display in LOESS curves pdf
new_regions_order <- c('VFC', 'DFC', 'MFC', 'OFC', 'M1C', 'S1C', 'IPC', 'STC',
                       'A1C', 'ITC', 'V1C', 'MD', 'AMY', 'HIP', 'STR', 'CBC')

# Use specified order to order the covid.gene.exp.long data frame
covid.gene.exp.long$Region <- factor(covid.gene.exp.long$Region,
                                     levels = new_regions_order)

# Create the plots and store them as loess.raw.stamp.collection
loess.raw.stamp.collection = ggplot(covid.gene.exp.long, 
                                    aes(x = Years, y = Expression)) +
  # Add in shading in the background to separate pre-natal, infancy, childhood, 
  # and adulthood
  annotate("rect", xmin = 0, xmax = 40/52, ymin = -Inf, ymax = Inf,
           alpha = 0.2, fill = "grey40") +
  annotate("rect", xmin = 40/52, xmax = 92/52, ymin = -Inf, ymax = Inf,
           alpha = 0.2, fill = "grey60") +
  annotate("rect", xmin = 92/52, xmax = 40/52 + 18, ymin = -Inf, ymax = Inf,
           alpha = 0.2, fill = "grey80") +
  # Add in a dashed line to separate pre and post-natal
  geom_vline(aes(xintercept = 40/52), linetype = "dashed") +
  # Add in the LOESS curves
  geom_smooth(method = "loess", span = 0.75,
              method.args = list(family = 'gaussian', degree = 2)) +
  # Scale the x-axis by log2 and overwrite the labels
  scale_x_continuous(trans = "log2", 
                     breaks = c(0, 10/52, 20/52, 40/52, 40/52 + 2/12,
                                40/52 + 6/12, 92/52,40/52 + 2, 40/52 + 4,
                                40/52 + 8, 40/52 + 16, 40/52 + 32,
                                40/52 + 64, 40/52 + 90),
                     labels = c("0 PCW", "10 PCW", "20 PCW", "0", "2 M",
                                "6 M", "1 Y", "2 Y", "4 Y", "8 Y", "16 Y",
                                "32 Y", "64 Y", "90 Y")) +
  # Separate the data into 16 panels (regions) for each gene
  facet_wrap(~Gene + Region) +
  # Label the axes
  labs(x = "Age", y = "Lifespan Trajectory of Gene Expression") +
  # Change up the theme for aesthetics and clarity
  theme_bw() +
  theme(axis.text.x = element_text(angle= 270, size = 7),
        axis.text.y = element_text(size = 7),
        axis.title.y = element_text(size = 9),
        panel.grid = element_blank(),
        strip.text.x = element_text(face = "bold"),
        strip.background = element_rect(fill = "grey75"))

# Determine how many pages we will need
num_pages <- length(genes)
```

```{r eval=FALSE, include=FALSE}
# Save the plots as a pdf
devEval(type = "pdf",
        path = here("Exploration"),
        name = "LOESS Curves of Raw Expression for all 29 Genes By Region with SE no points",
        height = 9, 
        width = 12,
        expr = {
          for (j in 1:num_pages) {
            # Plot LOESS curve for each and facet wrap by gene
              print(loess.raw.stamp.collection + 
                    facet_wrap_paginate(vars(Gene, Region), 
                                        ncol = 4, nrow = 4, 
                                        page = j, scales = "free_y"))
            }
        })
```

#### PDF of LOESS Curves (Raw/Non Z-scored) for 16 Regions by Gene 
```{r}
# Use for loop to create a separate PDF for each region, each PDF contains
# 29 LOESS plots for each gene. This will result in 16 PDFs which will later
# be merged together into one PDF.
for (i in new_regions_order) {

  # Create subset containing loess data for region i only
  subset <- subset(covid.gene.exp.long, Region == i)
  
  # Create the plots and store them as loess.raw.stamp.collection
  loess.raw.stamp.collection <- ggplot(subset, 
                                       aes(x = Years, y = Expression)) +
    # Add in shading in the background to separate pre-natal, infancy, childhood, 
    # and adulthood
    annotate("rect", xmin = 0, xmax = 40/52, ymin = -Inf, ymax = Inf,
             alpha = 0.2, fill = "grey40") +
    annotate("rect", xmin = 40/52, xmax = 92/52, ymin = -Inf, ymax = Inf,
             alpha = 0.2, fill = "grey60") +
    annotate("rect", xmin = 92/52, xmax = 40/52 + 18, ymin = -Inf, ymax = Inf,
             alpha = 0.2, fill = "grey80") +
    # Add in a dashed line to separate pre and post-natal
    geom_vline(aes(xintercept = 40/52), linetype = "dashed") +
    # Add in the LOESS curves
    geom_smooth(method = "loess", span = 0.75,
                method.args = list(family = 'gaussian', degree = 2)) +
    # Scale the x-axis by log2 and overwrite the labels
    scale_x_continuous(trans = "log2", 
                       breaks = c(0, 10/52, 20/52, 40/52, 40/52 + 2/12,
                                  40/52 + 6/12, 92/52, 40/52 + 2, 40/52 + 4,
                                  40/52 + 8, 40/52 + 16, 40/52 + 32,
                                  40/52 + 64, 40/52 + 90),
                       labels = c("0 PCW", "10 PCW", "20 PCW", "0", "2 M",
                                  "6 M", "1 Y", "2 Y", "4 Y", "8 Y", "16 Y",
                                  "32 Y", "64 Y", "90 Y")) +
    # Separate the data into 29 panels (genes) for each region
    facet_wrap(~Region + Gene) +
    # Label the axes
    labs(x = "Age", y = "Lifespan Trajectory of Gene Expression") +
    # Change up the theme for aesthetics and clarity
    theme_bw() +
    theme(axis.text.x = element_text(angle = 270, size = 7),
          axis.text.y = element_text(size = 7),
          axis.title.y = element_text(size = 9),
          panel.grid = element_blank(),
          strip.text.x = element_text(face = "bold"),
          strip.background = element_rect(fill = "grey75")) 
  
  # For each region, we need 2 pages
  # 16 genes on first page, 13 genes on second page
  num_pages2 <- 2
  
  # Save the plots as a pdf
  devEval(type = "pdf",
          path = here("Exploration"),
          name = paste0(
            "LOESS Curves of Raw Expression for ", i, " by Gene with SE no points"),
          height = 9, 
          width = 12,
          expr = {
            for (j in 1:num_pages2) {
              # Plot LOESS curve for each region and facet wrap by gene
                print(loess.raw.stamp.collection + 
                      facet_wrap_paginate(~Region + Gene,
                                          ncol = 4,
                                          nrow = 4,
                                          page = j,
                                          scales = "free_y"))
            }
          })
}
```

```{r, merge pdfs into one}
# Merge all 16 PDFs into one — paste path of each pdf and specify output path
pdf_combine(
  c("/Users/rachelkwan/Documents/Lab/summer_2021/Exploration/LOESS Curves of Raw Expression for VFC by Gene with SE no points.pdf",
    "/Users/rachelkwan/Documents/Lab/summer_2021/Exploration/LOESS Curves of Raw Expression for DFC by Gene with SE no points.pdf",
    "/Users/rachelkwan/Documents/Lab/summer_2021/Exploration/LOESS Curves of Raw Expression for MFC by Gene with SE no points.pdf",
   "/Users/rachelkwan/Documents/Lab/summer_2021/Exploration/LOESS Curves of Raw Expression for OFC by Gene with SE no points.pdf",
   "/Users/rachelkwan/Documents/Lab/summer_2021/Exploration/LOESS Curves of Raw Expression for M1C by Gene with SE no points.pdf",
   "/Users/rachelkwan/Documents/Lab/summer_2021/Exploration/LOESS Curves of Raw Expression for S1C by Gene with SE no points.pdf",
   "/Users/rachelkwan/Documents/Lab/summer_2021/Exploration/LOESS Curves of Raw Expression for IPC by Gene with SE no points.pdf",
   "/Users/rachelkwan/Documents/Lab/summer_2021/Exploration/LOESS Curves of Raw Expression for STC by Gene with SE no points.pdf",
   "/Users/rachelkwan/Documents/Lab/summer_2021/Exploration/LOESS Curves of Raw Expression for A1C by Gene with SE no points.pdf",
   "/Users/rachelkwan/Documents/Lab/summer_2021/Exploration/LOESS Curves of Raw Expression for ITC by Gene with SE no points.pdf",
   "/Users/rachelkwan/Documents/Lab/summer_2021/Exploration/LOESS Curves of Raw Expression for V1C by Gene with SE no points.pdf",
   "/Users/rachelkwan/Documents/Lab/summer_2021/Exploration/LOESS Curves of Raw Expression for MD by Gene with SE no points.pdf",
   "/Users/rachelkwan/Documents/Lab/summer_2021/Exploration/LOESS Curves of Raw Expression for AMY by Gene with SE no points.pdf",
   "/Users/rachelkwan/Documents/Lab/summer_2021/Exploration/LOESS Curves of Raw Expression for HIP by Gene with SE no points.pdf",
   "/Users/rachelkwan/Documents/Lab/summer_2021/Exploration/LOESS Curves of Raw Expression for STR by Gene with SE no points.pdf",
   "/Users/rachelkwan/Documents/Lab/summer_2021/Exploration/LOESS Curves of Raw Expression for CBC by Gene with SE no points.pdf"),
   output = "/Users/rachelkwan/Documents/Lab/summer_2021/Exploration/LOESS Curves of Raw Expression for all 16 Regions by Gene with SE no points.pdf")
```

# 2. LOESS Curves By Gene and By Region

## 2.1. LOESS Curves By Gene

## 2.2. LOESS Curves by Region

