---
title: "COVID-19 Project: LOESS Curves for 28 Genes and 16 Brain Regions"
author: "Rachel Kwan and Jonathan Zaslavsky"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include = FALSE}
library(knitr) # For formatting the document
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
```

## Relevant Packages
```{r}
library(tidyverse) # For ggplot2, dplyr
library(here) # To read in data from directory
library(reshape2) # For melt()
library(R.devices) # Ease of use for the pdf function
library(ggforce) # For the facet_wrap_paginate()
```

```{r, echo = FALSE, include = FALSE}
# Import dataset (already log2 transformed data) 
infl.gene.exp.raw = read_csv(here("Data", "COVID Data - 28 Genes - 1281 Samples from 16 Regions - No Transformation.csv"))
```

## Import and Prepare Dataset
```{r}
# Load full dataset
all_expression <- read_csv(here("Data", "COVID Data - 28 Genes - 1281 Samples from 16 Regions - No Transformation.csv"))

# Change dataset into long format
all_expression.long <- pivot_longer(all_expression, cols = -c("Sample",
                                                              "Region",
                                                              "Years"),
                                    names_to = "Gene",
                                    values_to = "Expression")

# Create a vector containing all of the names of the genes of interest.
genes <- as_vector(colnames(all_expression)[4:31])

# Define a vector containing the 16 brain regions of interest
# (ordered alphabetically), for later use.
regions <- as_vector(unique(all_expression$Region))
```

## Z-score Normalize the Gene Expression
```{r z-score the data}
covid.gene.exp.numeric = all_expression %>%
  #Make sure all expression values are numeric in R
  mutate_at(genes, as.numeric)

covid.gene.exp.zscore = covid.gene.exp.numeric %>%
  # Group the data by region
  group_by(Region) %>%
  # 'Nest' the data (i.e. build data frames for each region inside of the data
  # frame)
  nest() %>%
  # z-score the data separately for each gene and region
  mutate(data = map(.x = data, .f = ~.x %>% mutate_at(genes, scale))) %>%
  # Unnest the data
  unnest()

# Change dataset into long format
covid.gene.exp.zscore.long <- pivot_longer(covid.gene.exp.zscore, 
                                           cols = -c("Sample",
                                                     "Region",
                                                     "Years"),
                                           names_to = "Gene",
                                           values_to = "Expression")
```

## Fit LOESS Curves for Z-score Data
```{r fit loess curves for z-scored data}
covid.gene.exp.zscore.loess <- covid.gene.exp.zscore %>%
  # Use the melt function to put the data in long format
  melt(id.vars = c("Sample", "Region", "Years"),
       variable.name = "Gene",
       value.name = "Expression") %>%
  # Group by region and Gene, then nest the data
  group_by(Region, Gene) %>%
  nest() %>%
  # Fit a LOESS model for each gene and region
  mutate(loess.model = map(.x = data, ~loess(data = .x, formula = Expression ~ log2(Years)))) %>%
  # Find the min and max age value for each curve
  mutate(min = map(.x = data, .f = ~min(log2(.x$Years))),
         max = map(.x = data, .f = ~max(log2(.x$Years)))) %>%
  unnest(min) %>% unnest(max) %>%
  # Create a vector of 80 log2 age values for each curve
  mutate(age.values = map2(.x = min, .y = max, ~seq(.x, .y, (.y - .x) / 79))) %>%
  # Extract the fitted LOESS expression values along this curve
  mutate(fitted.values = map2(.x = loess.model, .y = age.values,
                              ~data.frame("Years" = 2^(.y), 
                                          "Expression" = predict(.x, newdata = .y))))

# Create a data frame including only the fitted values of the LOESS curves of
# the z-scored data
covid.gene.exp.zscore.loess.fitted = covid.gene.exp.zscore.loess %>%
  # Select necessary columns
  select(Region, Gene, fitted.values) %>%
  # Unnest the fitted values
  unnest(cols = c(fitted.values))
```

## Plot a Test LOESS Curve

Raw (log2 transformed) data
```{r}
# Representative LOESS curve using raw (and log2 transformed) data
test.raw <- all_expression.long %>% 
  # Filter for only DPP4 expression values from the V1C
  filter(Gene == "ACE2", Region == "V1C")

ggplot(test.raw, aes(x = Years, y = Expression)) +
  # Add in shading in the background to separate pre-natal, infancy, childhood, 
  # and adulthood
  annotate("rect", xmin = 0, xmax = 40/52, ymin = -Inf, ymax = Inf, 
           alpha = 0.2, fill = "grey40") +
  annotate("rect", xmin = 40/52, xmax = 92/52, ymin = -Inf, ymax = Inf, 
           alpha = 0.2, fill = "grey60") +
  annotate("rect", xmin = 92/52, xmax = 40/52 + 18, ymin = -Inf, ymax = Inf, 
           alpha = 0.2, fill = "grey80") +
  # Add in a dashed line to separate pre- and post-natal
  geom_vline(aes(xintercept = 40/52), linetype = "dashed") +
  # Add in the LOESS curves
  geom_smooth(method = "loess", span = 0.7091, 
              method.args = list(family = 'gaussian', degree = 2)) +
  # Add points
  geom_point(size = 1) +
  # Scale the x-axis by log2 and overwrite the labels
  scale_x_continuous(trans = "log2", 
                     breaks = c(0, 10/52, 20/52, 40/52, 40/52 + 2/12,
                                40/52 + 6/12, 92/52, 40/52 + 2, 40/52 + 4,
                                40/52 + 8, 40/52 + 16, 40/52 + 32,40/52 + 64,
                                40/52 + 90),
                     labels = c("0 PCW", "10 PCW", "20 PCW", "0", 
                                "2 M", "6 M", "1 Y", "2 Y", "4 Y", "8 Y",
                                "16 Y", "32 Y", "64 Y", "90 Y")) +
  # Separate the data into 16 panels for each region
  facet_wrap(~Gene + Region) +
  # Label the axes
  labs(x = "Age", 
       y = "Lifespan Trajectory of Gene Expression") +
  # Change up the theme for aesthetics and clarity
  theme_bw() +
  theme(axis.text.x = element_text(angle = 270, size = 7),
        axis.text.y = element_text(size = 7),
        axis.title.y = element_text(size = 9),
        panel.grid = element_blank())

```

Z-score (and log2 transformed) data
```{r}
# Representative LOESS curve using z-score (and log2 transformed) data
test.zscore <- covid.gene.exp.zscore.long %>% 
  # Filter for only ACE2 expression values from the V1C
  filter(Gene == "ACE2", Region == "V1C")

ggplot(test.zscore, aes(x = Years, y = Expression)) +
  # Add in shading in the background to separate pre-natal, infancy, childhood,
  # and adulthood
  annotate("rect", xmin = 0, xmax = 40/52, ymin = -Inf, ymax = Inf,
           alpha = 0.2, fill = "grey40") +
  annotate("rect", xmin = 40/52, xmax = 92/52, ymin = -Inf, ymax = Inf,
           alpha = 0.2, fill = "grey60") +
  annotate("rect", xmin = 92/52, xmax = 40/52 + 18, ymin = -Inf, ymax = Inf,
           alpha = 0.2, fill = "grey80") +
  # Add in a dashed line to separate pre and post-natal
  geom_vline(aes(xintercept = 40/52), linetype = "dashed") +
  # Add in the LOESS curves
  geom_smooth(method = "loess", span = 0.7091,
              method.args = list(family = 'gaussian', degree = 2)) +
  # Add points
  geom_point(size = 1) +
  # Scale the x-axis by log2 and overwrite the labels
  scale_x_continuous(trans = "log2", 
                     breaks = c(0, 10/52, 20/52, 40/52, 40/52 + 2/12,
                                40/52 + 6/12, 92/52, 40/52 + 2, 40/52 + 4,
                                40/52 + 8, 40/52 + 16, 40/52 + 32, 40/52 + 64,
                                40/52 + 90),
                    labels = c("0 PCW", "10 PCW", "20 PCW", "0",
                               "2 M", "6 M", "1 Y", "2 Y", "4 Y", "8 Y",
                               "16 Y", "32 Y", "64 Y", "90 Y")) +
  # Separate the data into 16 panels for each region
  facet_wrap(~Gene + Region) +
  # Label the axes
  labs(x = "Age", y = "Lifespan Trajectory of Gene Expression") +
  # Change up the theme for aesthetics and clarity
  theme_bw() +
  theme(axis.text.x = element_text(angle = 270, size = 7),
        axis.text.y = element_text(size = 7),
        axis.title.y = element_text(size = 9),
        panel.grid = element_blank())
```


## Plot the LOESS Curves and Save as PDFs
```{r, include = FALSE}
output_path = here("./Plots/LOESS Curves")
```

### Prepare PDF of LOESS curves for z-score data with 28 genes by region
```{r}
# Create the plots and store them as loess.raw.stamp.collection
loess.raw.zscore.stamp.collection <- ggplot(covid.gene.exp.zscore.long, 
                                           aes(x = Years, y = Expression)) +
  # Add in shading in the background to separate pre-natal, infancy, childhood,
  # and adulthood
  annotate("rect", xmin = 0, xmax = 40/52, ymin = -Inf, ymax = Inf,
           alpha = 0.2, fill = "grey40") +
  annotate("rect", xmin = 40/52, xmax = 92/52, ymin = -Inf, ymax = Inf,
           alpha = 0.2, fill = "grey60") +
  annotate("rect", xmin = 92/52, xmax = 40/52 + 18, ymin = -Inf, ymax = Inf,
           alpha = 0.2, fill = "grey80") +
  # Add in a dashed line to separate pre and post-natal
  geom_vline(aes(xintercept = 40/52), linetype = "dashed") +
  # Add in the LOESS curves
  geom_smooth(method = "loess", span = 0.7091,
              method.args = list(family = 'gaussian', degree = 2)) +
  # Add points
  # geom_point(size = 1) +
  # Scale the x-axis by log2 and overwrite the labels
  scale_x_continuous(trans = "log2", 
                    breaks = c(0, 10/52, 20/52, 40/52, 40/52 + 2/12,
                               40/52 + 6/12, 92/52, 40/52 + 2, 40/52 + 4,
                               40/52 + 8, 40/52 + 16, 40/52 + 32, 40/52 + 64,
                               40/52 + 90),
                    labels = c("0 PCW", "10 PCW", "20 PCW", "0",
                               "2 M", "6 M", "1 Y", "2 Y", "4 Y", "8 Y",
                               "16 Y", "32 Y", "64 Y", "90 Y")) +
  # Separate the data into 16 panels for each region
  facet_wrap(~Gene + Region) +
  # Label the axes
  labs(x = "Age", y = "Lifespan Trajectory of Gene Expression") +
  # Change up the theme for aesthetics and clarity
  theme_bw() +
  theme(axis.text.x = element_text(angle = 270, size = 7),
        axis.text.y = element_text(size = 7),
        axis.title.y = element_text(size = 9),
        panel.grid = element_blank())

# Determine how many pages we will need
num_pages = length(genes)

# Save the plots as a PDF
devEval(type = "pdf",
        path = here(output_path),
        name = "LOESS Curves of Non-Transformed Z-scored Expression for all 28 Genes by Region with SE no points",
        height = 9, 
        width = 9,
        expr = {
          
          for (j in 1:num_pages) {
            # Plot LOESS curve for each gene and facet wrap by region
              print(loess.raw.zscore.stamp.collection + 
                      facet_wrap_paginate(vars(Gene, Region),
                                          ncol = 4,
                                          nrow = 4,
                                          page = j,
                                          scales = "free_y"))
            }

        })
```

### Prepare PDF of LOESS curves for z-score data with 16 brain regions by genes
```{r}
# Create the plots and store them as loess.raw.zscore.stamp.collection.v2
loess.raw.zscore.stamp.collection.v2 <- ggplot(covid.gene.exp.zscore.long, 
                                           aes(x = Years, y = Expression)) +
  # Add in shading in the background to separate pre-natal, infancy, childhood,
  # and adulthood
  annotate("rect", xmin = 0, xmax = 40/52, ymin = -Inf, ymax = Inf,
           alpha = 0.2, fill = "grey40") +
  annotate("rect", xmin = 40/52, xmax = 92/52, ymin = -Inf, ymax = Inf,
           alpha = 0.2, fill = "grey60") +
  annotate("rect", xmin = 92/52, xmax = 40/52 + 18, ymin = -Inf, ymax = Inf,
           alpha = 0.2, fill = "grey80") +
  # Add in a dashed line to separate pre and post-natal
  geom_vline(aes(xintercept = 40/52), linetype = "dashed") +
  # Add in the LOESS curves
  geom_smooth(method = "loess", span = 0.7091,
              method.args = list(family = 'gaussian', degree = 2)) +
  # Add points
  #geom_point(size = 1) +
  # Scale the x-axis by log2 and overwrite the labels
  scale_x_continuous(trans = "log2", 
                     breaks = c(0, 10/52, 20/52, 40/52, 40/52 + 2/12,
                                40/52 + 6/12, 92/52, 40/52 + 2, 40/52 + 4,
                                40/52 + 8, 40/52 + 16, 40/52 + 32, 40/52 + 64,
                                40/52 + 90),
                     labels = c("0 PCW", "10 PCW", "20 PCW", "0",
                                "2 M", "6 M", "1 Y", "2 Y", "4 Y", "8 Y",
                                "16 Y", "32 Y", "64 Y", "90 Y")) +
  # Separate the data into 16 panels for each region
  facet_wrap(~Region + Gene) +
  # Label the axes
  labs(x = "Age",
       y = "Lifespan Trajectory of Gene Expression") +
  # Change up the theme for aesthetics and clarity
  theme_bw() +
  theme(axis.text.x = element_text(angle = 270, size = 7),
        axis.text.y = element_text(size = 7),
        axis.title.y = element_text(size = 9),
        panel.grid = element_blank()) +
  facet_wrap_paginate(vars(Region, Gene), ncol = 7, nrow = 4, scales = "free_y")

# Determine how many pages we will need; multiply by two because each region
# needs two pages
num_pages = 2 * length(regions)

num_pages2 <- n_pages(loess.raw.zscore.stamp.collection.v2)

# Save the plots as a PDF
devEval(type = "pdf",
        path = here(output_path),
        name = "LOESS Curves of Non-Transformed Z-scored Expression for all 16 Regions by Gene with SE no points",
        height = 11, 
        width = 20,
        expr = {
          for (j in 1:num_pages2) {
            # Plot LOESS curve for each region and facet wrap by gene
              print(loess.raw.zscore.stamp.collection.v2 + 
                    facet_wrap_paginate(vars(Region, Gene),
                                        ncol = 7,
                                        nrow = 4,
                                        page = j,
                                        scales = "free_y"))
          }
          
        })
```
